---
title: "Viz II"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(patchwork)
library(haven)
```
library(tidyverse)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


# Import the weather data

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USW00022534", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2021-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = case_match(
      id, 
      "USW00094728" ~ "CentralPark_NY", 
      "USW00022534" ~ "Molokai_HI",
      "USS0023B17S" ~ "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())
```


# Make a scatterplot but fancy this time.

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = .3) + 
  labs(
    title = "Temperature scatterplot",
    x = "Minimum Temp (C)",
    y = "Maximum Temp (C)",
    color = "Location", 
    caption = "Weather data taken from rnoaa package for three stations."
  )
```

## Scales -- start with `x` and `y` and then do `color`.

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = .3) + 
  labs(
    title = "Temperature scatterplot",
    x = "Minimum Temp (C)",
    y = "Maximum Temp (C)",
    color = "Location", 
    caption = "Weather data taken from rnoaa package for three stations."
  ) + 
  scale_x_continuous(
    breaks = c(-15, 0, 20),
    labels = c("-15C", "0", "20")
  ) + 
  scale_y_continuous(
    limits = c(0, 30),
    transform = "sqrt"
  )
```


## Look at color:

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = .3) + 
  labs(
    title = "Temperature scatterplot",
    x = "Minimum Temp (C)",
    y = "Maximum Temp (C)",
    color = "Location", 
    caption = "Weather data taken from rnoaa package for three stations."
  ) + 
  viridis::scale_color_viridis(discrete = TRUE) # fancy color
```



## Themes


```{r}
ggp_scatterplot = 
  weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = .3) + 
  labs(
    title = "Temperature scatterplot",
    x = "Minimum Temp (C)",
    y = "Maximum Temp (C)",
    color = "Location", 
    caption = "Weather data taken from rnoaa package for three stations."
  ) + 
  viridis::scale_color_viridis(discrete = TRUE)
```

```{r}
ggp_scatterplot +  # 指代上面的所有内容
  theme(legend.position = "bottom") # 图标放在下面
```

```{r}
ggp_scatterplot + 
  theme_bw() +  #背景black an white, 没有条纹了
  theme(legend.position = "bottom")
```

Order matters ...

```{r}
ggp_scatterplot + 
  theme(legend.position = "bottom") + 
  theme_bw() 
```

```{r}
ggp_scatterplot + 
  theme(legend.position = "bottom") + 
  theme_minimal() 
```


```{r}
ggp_scatterplot + 
  theme(legend.position = "bottom") + 
  theme_classic() 
```


```{r}
ggp_scatterplot + 
  theme(legend.position = "bottom") + 
  ggthemes::theme_excel() 
```


New scatterplot

```{r}
weather_df |> 
  ggplot(aes(x = date, y = tmax, color = name, size = prcp)) + 
  geom_point(alpha = .3) + 
  geom_smooth(se = FALSE) + 
  viridis::scale_color_viridis(discrete = TRUE) + 
  labs(
    x = "Date", 
    y = "Maximum Temp (C)",
    title = "Seasonal variation in Max Temp",
    size = "Precipitation"
  ) + 
  theme_minimal() + 
  theme(legend.position = "bottom")
```



Extra bonus stuff in `ggplot`

Use different datasets in different `geom`s

```{r}
central_park_df = 
  weather_df |> 
  filter(name == "CentralPark_NY")

molokai_df = 
  weather_df |> 
  filter(name == "Molokai_HI")

molokai_df |> 
  ggplot(aes(x = date, y = tmax, color = name)) + 
  geom_point() +                        # scatter points for molokai_df
  geom_line(data = central_park_df)  #line for central_park_df
```




## Multiple panels

### Category variable 每一个类别level分开展示
```{r}
weather_df |> 
  ggplot(aes(x = tmax, fill = name)) + 
  geom_density() + 
  facet_grid(. ~ name)
```

### 将几个图拼在一起
```{r}
ggp_tmax_tmin = 
  weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = .3)

ggp_tmax_density = 
  weather_df |> 
  ggplot(aes(x = tmax, fill = name)) + 
  geom_density(alpha = .3) 

ggp_tmax_date = 
  weather_df |> 
  ggplot(aes(x = date, y = tmax, color = name)) + 
  geom_point() + 
  geom_smooth(se = FALSE)

# 设置这几个图的位置:
(ggp_tmax_tmin + ggp_tmax_density) / ggp_tmax_date
```


## Data manipulation
给category排序
```{r}
weather_df |> 
  mutate(name = fct_relevel(name, c("Molokai_HI", "CentralPark_NY", "Waterhole_WA"))) |>  # 给category排序 
  ggplot(aes(x = name, y = tmax, fill = name)) + 
  geom_violin(alpha = .5)
```



PULSE data next. 


```{r}
pulse_df = 
  read_sas("data/public_pulse_data.sas7bdat") |> 
  janitor::clean_names() |> 
#将4个variable 变成一个长的 2个variable(一列是4个category, 一列是对应的值)
  pivot_longer(
    cols = bdi_score_bl:bdi_score_12m,
    names_to = "visit", 
    values_to = "bdi_score", 
    names_prefix = "bdi_score_"
  ) |> 
  mutate(visit = ifelse(visit == "bl", "00m", visit))

pulse_df |> 
  ggplot(aes(x = visit, y = bdi_score)) + 
  geom_boxplot()
```



Make an FAS plot.


```{r}
# 拆成2列, 以根据其中一个数字作为key组成两张表
litters_df = 
  read_csv("data/FAS_litters.csv", na = c("NA", ".", "")) |> 
  janitor::clean_names() |> 
  separate(group, into = c("dose", "tx_day"), 3) Con7分开成“Con”和“7”
# tidy data
pups_df = 
  read_csv("data/FAS_pups.csv", na = c("NA", ".", "")) |> 
  janitor::clean_names() |> 
  pivot_longer(
    pd_ears:pd_walk,
    names_to = "outcome", 
    values_to = "pn_day",
    names_prefix = "pd_"
  )
# 组合两张表  
fas_df = 
  left_join(pups_df, litters_df, by = "litter_number")

fas_df |> 
  drop_na(tx_day) |> 
  ggplot(aes(x = dose, y = pn_day)) +    #每一个小图的x, y
  geom_boxplot() + 
  facet_grid(tx_day ~ outcome)           #一整张大图的x, y
```














